### 2017-04-29 paper/jfp-submission/evaluation.tex
add threats for localization

\subsection{Methodology}
\label{sec:methodology}
We answer the first two questions on two sets of ill-typed programs,
\ie\ programs that were rejected by the \ocaml\ compiler because of a
type error.
%
The first dataset comes from the Spring 2014 undergraduate Programming
Languages (CSE 130) course at UC San Diego.
%
We recorded each interaction with the \ocaml\ top-level system over the
course of the first three assignments (IRB
% \# hidden for blind review),
\#140608),
from which we extracted \ucsdsize\ distinct, ill-typed \ocaml\ programs.
%
The second dataset --- widely used in the literature --- comes from a
graduate-level course at the University of Washington~\cite{Lerner2006-pj},
from which we extracted 284 ill-typed programs.
%
Both datasets contain relatively small programs, the largest being 348
SLoC; however, they demonstrate a variety of functional programming
idioms including (tail) recursive functions, higher-order functions,
polymorphic and algebraic data types. % and expression evaluators.

@@

\paragraph{Example: Recursion with Bad Operator}
The recursive function @sqsum@ should square each
element of the input list and then compute the sum
of the result.
%
\begin{ecode}
  let rec sqsum xs = match xs with
    | [] -> 0
    | h::t -> (*@\hlOcaml{\hlSherrloc{sqsum t}}@*) @ (h * h)
\end{ecode}
%
Unfortunately the student has used the list-append
operator |@| instead of \texttt{+}. % to compute the sum.
%
Both \ocaml\ and \sherrloc\ blame the \emph{wrong location},
the recursive call @sqsum t@, with the message
%
\begin{verbatim}
  This expression has type
    int
  but an expression was expected of type
    'a list
\end{verbatim}
%
\begin{center}
  \includegraphics[height=125px]{sqsum.png}
\end{center}
%
\toolname\ produces the above trace showing how the evaluation of
@sqsum [1]@ gets stuck.
%
The trace highlights the entire stuck term
(not just the recursive call), emphasizing
the \emph{conflict} between @int@ and @list@
rather than assuming one or the other is correct.

>>>>

\subsection{Methodology}
\label{sec:methodology}
We answer the first two questions on two sets of ill-typed programs,
\ie\ programs that were rejected by the \ocaml\ compiler because of a
type error.
%
The first dataset comes from the Spring 2014 undergraduate Programming
Languages (CSE 130) course at UC San Diego.
%
We recorded each interaction with the \ocaml\ top-level system over the
course of the first three assignments (IRB
% \# hidden for blind review),
\#140608),
from which we extracted \ucsdsize\ distinct, ill-typed \ocaml\ programs
from a cohort of 46 students..
%
The second dataset --- widely used in the literature --- comes from a
graduate-level course at the University of Washington~\cite{Lerner2006-pj},
from which we extracted 284 ill-typed programs.
%
Both datasets contain relatively small programs, the largest being 348
SLoC; however, they demonstrate a variety of functional programming
idioms including (tail) recursive functions, higher-order functions,
polymorphic and algebraic data types. % and expression evaluators.

@@

\paragraph{Example: Recursion with Bad Operator}
The recursive function @sqsum@ should square each
element of the input list and then compute the sum
of the result.
%
\begin{ecode}
  let rec sqsum xs = match xs with
    | [] -> 0
    | h::t -> (*@\hlOcaml{\hlSherrloc{sqsum t}}@*) @ (h * h)
\end{ecode}
%
Unfortunately the student has used the list-append
operator |@| instead of \texttt{+}. % to compute the sum.
%
Both \ocaml\ and \sherrloc\ blame the \emph{wrong location},
the recursive call @sqsum t@, with the message
%
\begin{verbatim}
  This expression has type
    int
  but an expression was expected of type
    'a list
\end{verbatim}
%
\toolname\ produces a trace showing how the evaluation of
@sqsum [1]@ gets stuck.
%
\begin{center}
  \includegraphics[height=125px]{sqsum.png}
\end{center}
%
The trace highlights the entire stuck term
(not just the recursive call), emphasizing
the \emph{conflict} between @int@ and @list@
rather than assuming one or the other is correct.

### 2017-04-29 paper/jfp-submission/locating.tex
add threats for localization

\ES{TODO: threats to validity}

>>>>

\paragraph{Threats to Validity}
Our benchmarks were drawn from students in an undergraduate course
at \ucsdbench\ and may not be representative of other student bodies.
%
We mitigate this threat with a large empirical evaluation of 1,569
programs, drawn from a cohort of 46 students.
%
A similar threat is that students are not industrial programmers, thus
our results may not translate to large-scale software engineering.
%
In particular, \toolname's reliance on running the program suggests that
scaling it to larger programs would be more difficult than static
methods.
%
However, in our experience programmers are able to construct a mental
model of type systems after sufficient exposure, at which point
traditional error reports may suffice.
%
We are thus particularly interested in aiding novice programmers as
they learn to work with the type system.

Our definition of the next well-typed program as the intended ground
truth answer is another threat to validity. Students might submit
multiple well-typed ``rewrites'' between the initial ill-typed program
and the final intended answer.
%
Our approach to discarding outliers is intended to mitigate this threat.
%
A similar threat is our removal of programs where any of the tools could
not produce an answer.
%
It may be, for example, that \mycroft and \sherrloc are particularly
effective on programs that do not admit dynamic witnesses.
%
Finally, our use of student fixes as oracles for the source of type
errors assumes that students are able to correctly identify the source.
%
As the students are in the process of learning \ocaml and the type
system, this assumption may be faulty, \emph{expert} users may disagree
with the student fixes.
%
We believe, however, that it is reasonable to use student fixes as
oracles, as the student is the best judge of what she \emph{intended} to
do.
